---
title: "Untitled"
author: "Eunice Chidiebube Egeonu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
        smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, error = F)
```

## Tools for Analysis

```{r}
library(tidyverse)
library(lubridate)
library(showtext)
library(ggtext)
library(cohorts)
library(arules)
library(arulesViz)
library(data.table)

font_add_google(name = "Lato", family = "Lato")

theme_set(
  theme_minimal() +
    theme(text = element_text(family = "Lato"))
)

```

## Data Loading and Cleaning

```{r}
data_path <- fs::dir_ls("data/")

load_and_clean <- function(x) {
  vroom::vroom(x) %>% ## Loads data fast in R
    janitor::clean_names() %>% ## Converts column name to snake case (i.e 'Customer ID' = customer_id)
    mutate(across(where(is.character), factor))
}

data <- map(data_path, load_and_clean)
```

Transaction table: We don't need records showing 0 sales or 0 quantity

```{r}
tranx_tbl <- data$`data/transaction_data.csv` %>%
  filter(sales_value > 0)

demographics_tbl <- data$`data/hh_demographic.csv`

coupon_redemption_tbl <- data$`data/coupon_redempt.csv`

coupon_tbl <- data$`data/coupon.csv`

campaign_desc_tbl <- data$`data/campaign_desc.csv`

campaign_table_tbl <- data$`data/campaign_table.csv`

product_tbl <- data$`data/product.csv`
```

## Gather data

### Coupon + Sales

```{r}
coupon_sales_tbl <-
  coupon_tbl %>% ## Dataframe with coupon info
  inner_join(product_tbl, by = "product_id") %>% # Get Details of products that campaign were ran on
  inner_join(tranx_tbl, by = "product_id") %>% # Get information of the sales on those campaigns
  inner_join(coupon_redemption_tbl,
    by = c("household_key", "day", "coupon_upc", "campaign")
  ) %>% ## Streamline to customers who redeemed those coupons
  filter(coupon_disc != 0 | retail_disc != 0)
```

### Data for Cohort Analysis

```{r}
## For cohort analysis, we just need the customer ID and date of purchase
## For data: We define our date to start from "2012/01/01" -> Day 1 just for the purpose of understanding customer retention. It doesn't mean the store started on that said date
cohort_tbl <-
  tranx_tbl %>%
  select(household_key, day) %>%
  mutate(date = ymd("2012/01/01") + days(day - 1), .before = household_key) %>%
  select(-day)
```

## Custom Plots

### Duration of Campaigns

```{r}
plot_1 <-
  campaign_desc_tbl %>%
  mutate(
    duration = end_day - start_day,
    campaign = paste0("Campaign-", campaign) %>%
      fct_reorder(duration)
  ) %>%
  select(-ends_with("_day")) %>%
  ggplot(aes(duration, campaign, col = description, fill = description)) +
  geom_segment(aes(x = 0, xend = duration, yend = campaign),
    lineend = "round",
    lwd = 6
  ) +
  geom_text(aes(label = duration),
    col = "white", family = "Quicksand",
    hjust = 1, fontface = "bold"
  ) +
  labs(
    title = "Duration of Campaigns",
    subtitle = "Campaigns in <span style = 'color:#0e7c7b;'>**TypeC**</span> usually run for longer periods while <span style = 'color:#918450;'>**TypeB**</span> campaign runs for shorter periods",
    x = "Duration (days)",
    y = "",
    fill = "",
    col = ""
  ) +
  scale_color_manual(values = c(
    "TypeA" = "#ffb563",
    "TypeB" = "#918450",
    "TypeC" = "#0e7c7b"
  )) +
  theme(
    plot.subtitle = element_markdown(),
    plot.title = element_text(size = 28, face = "bold"),
    plot.title.position = "plot",
    legend.position = "top",
    legend.justification = "right"
  )

plot_1
```

### Coupon Redemption by Income Groups
```{r}
coupon_usage_by_age <-
  coupon_redemption_tbl %>%
  distinct(household_key) %>%
  inner_join(demographics_tbl) %>%
  count(income_desc, name = "total_c") %>%
  left_join(demographics_tbl %>%
    count(income_desc, name = "total")) %>%
  mutate(perc = total_c / total)

coupon_usage_by_age$income_desc <-
  factor(
    coupon_usage_by_age$income_desc,
    levels = c(
      "Under 15K",
      "15-24K",
      "25-34K",
      "35-49K",
      "50-74K",
      "75-99K",
      "100-124K",
      "125-149K",
      "150-174K",
      "175-199K",
      "200-249K",
      "250K+"
    ),
    ordered = T
  )

plot_4 <-
  coupon_usage_by_age %>%
  ggplot(aes(perc, income_desc)) +
  geom_vline(
    xintercept = 0.30,
    size = 1.2,
    col = "red"
  ) +
  geom_vline(
    xintercept = 0.5,
    size = 1.1,
    col = "yellow"
  ) +
  geom_vline(
    xintercept = 0.75,
    size = 1.0,
    col = "green"
  ) +
  geom_segment(
    aes(x = 0, xend = 1, yend = income_desc),
    lwd = 3.7,
    lineend = "round",
    col = "#2b2d42"
  ) +
  geom_segment(
    aes(x = 0, xend = perc, yend = income_desc),
    lwd = 2.5,
    lineend = "round",
    col = "white"
  ) +
  coord_flip() +
  scale_x_continuous(labels = scales::percent) +
  scale_y_discrete(expand = expansion(add = c(.5, 1))) +
  annotate(
    geom = "text",
    x = 0.78,
    y = 12.5,
    label = "75%",
    family = "quicksand",
    size = 4,
    fontface = "bold"
  ) +
  annotate(
    geom = "text",
    x = 0.53,
    y = 12.5,
    label = "50%",
    family = "quicksand",
    size = 4,
    fontface = "bold"
  ) +
  annotate(
    geom = "text",
    x = 0.33,
    y = 12.5,
    label = "30%",
    family = "quicksand",
    size = 4,
    fontface = "bold"
  ) +
  labs(
    title = "Coupon Redemption by Income Groups",
    subtitle = "The plot below shows the percentage coupon redemption by Income Groups",
    x = NULL,
    y = NULL,
  ) +
  theme(
    axis.text.y = element_blank(),
    plot.title = element_text(size = 28),
    plot.title.position = "plot"
  )

plot_4
```

### Departments with the most redeemed coupons
```{r}
donut_chart_tbl <-
  coupon_sales_tbl %>%
  select(department) %>%
  mutate(dept = fct_lump_prop(department, 0.05, other_level = "Others")) %>%
  count(dept) %>%
  mutate(ymax = cumsum(n))

donut_chart_tbl$ymin <- c(0, head(donut_chart_tbl$ymax, n = -1))

donut_chart_tbl$labelPosition <-
  (donut_chart_tbl$ymax + donut_chart_tbl$ymin) / 2

donut_chart_tbl$label <-
  paste0(donut_chart_tbl$dept, "\n value: ", donut_chart_tbl$n)

plot_5 <-
  donut_chart_tbl %>%
  ggplot(aes(
    ymax = ymax,
    ymin = ymin,
    xmax = 4,
    xmin = 3,
    fill = dept
  )) +
  geom_rect(col = "grey10") +
  # scale_fill_brewer(palette = 4) +
  coord_polar(theta = "y") +
  xlim(c(1, 4)) +
  labs(
    title = "Departments with the most redeemed coupons",
    subtitle = "Over 70% of coupons are used on Groceries",
    fill = NULL
  ) +
  scale_fill_manual(values = c("#495867", "#2a9d8f", "#e9c46a", "#e9c46a", "#f4a261")) +
  theme_void() +
  theme(
    text = element_text(family = "Lato"),
    legend.position = c(0.53, 0.5),
    legend.text.align = 0,
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5)
  )

plot_5
```


## Correlation

In here, we are trying to see if there's any correlation between the duration of a campaign and the quantity of purchases.

```{r}
sample_tbl <-
  coupon_sales_tbl %>%
  select(campaign, quantity, sales_value)

get_total_quantity <- function(x) {
  sample_tbl %>%
    filter(campaign == x) %>%
    summarize(total_quantity = sum(quantity)) %>%
    pull(total_quantity)
}
```

```{r}
duration_quantity <-
  campaign_desc_tbl %>%
  mutate(duration = end_day - start_day) %>%
  select(-ends_with("_day")) %>%
  mutate(total_quantity = map_dbl(campaign, get_total_quantity))
```

With a correlation of 0.032 and a p-value of 0.8632, there's isn't any correlation between length of campaign and quantity of products bought

```{r}
cor.test(duration_quantity$duration, duration_quantity$total_quantity)
```

## Cohort Analysis

```{r}
tranx_data_long <-
  cohort_tbl %>%
  cohort_table_month(household_key, date) %>%
  shift_left_pct() %>%
  pivot_longer(-cohort) %>%
  mutate(time = as.numeric(str_remove(name, "t")))

plot_2 <-
  tranx_data_long %>%
  filter(value > 0, cohort <= 7, time > 0) %>%
  mutate(line_type = ifelse(cohort %in% 1:4, "yes", "no")) %>%
  ggplot(aes(time, value, colour = factor(cohort), group = cohort)) +
  geom_line(aes(lty = line_type), alpha = 0.7, size = 0.9) +
  labs(
    title = "Cohort Analysis",
    subtitle = "The retention rate for the first 4 cohorts are very high",
    col = "Cohort",
    y = "Retention rate",
    x = "Time (months)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_linetype_manual(values = c(3, 1)) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 22)
  ) +
  guides(lty = "none")

plot_2
```

## Association Rules Mining

Preparing data for market basket analysis
```{r}
## Get product name
# transaction_tbl <-
#   tranx_tbl %>%
#   left_join(product_tbl %>%
#               select(product_id, sub_commodity_desc), by = "product_id")
#
# ## Converting to a transaction dataset
# purchased_items <- dcast(
#   setDT(
#     transaction_tbl %>%
#       group_by(basket_id) %>%
#       select(basket_id, sub_commodity_desc) %>%
#       distinct(sub_commodity_desc, .keep_all = TRUE)
#   ),
#   basket_id ~ rowid(basket_id)
# ) %>%
#   select(!basket_id)
#
# ## Saving data
# write.csv(
#   purchased_items,
#   "purchased_items.csv",
#   quote = FALSE,
#   row.names = FALSE,
#   col.names = FALSE,
#   na =  ""
# )

## Loading data
transaction <-
  read.transactions("purchased_items.csv",
    format = "basket",
    sep = ","
  )
```

```{r}
itemFrequencyPlot(transaction,
  topN = 25, type = "relative",
  main = "Most Frequently Purchased Items",
  ylab = "Relative Item Frequency", cex = 0.6
)
```

```{r}
plot_3 <-
  coupon_sales_tbl %>%
  count(sub_commodity_desc) %>%
  mutate(perc = n / sum(n)) %>%
  slice_max(order_by = perc, n = 20) %>%
  mutate(sub_commodity_desc = fct_reorder(sub_commodity_desc, perc)) %>%
  ggplot(aes(perc, sub_commodity_desc)) +
  geom_segment(aes(x = 0, xend = perc, yend = sub_commodity_desc),
    lineend = "butt"
  ) +
  geom_point(aes(size = perc), col = "#283618", show.legend = F) +
  labs(
    title = "Most frequently redeemed coupons",
    subtitle = "About **72%** of the coupons redeemed were on Groceries, here are the top 20 products the copuons were redeemed on",
    x = "Relative Coupon Use",
    y = "",
    fill = "",
    col = ""
  ) +
  scale_x_continuous(labels = scales::percent) +
  theme(
    plot.subtitle = element_markdown(),
    plot.title = element_text(size = 28),
    plot.title.position = "plot"
  )

plot_3
```

Having a strict set of rules
```{r}
rules <-
  apriori(transaction,
    parameter = list(
      supp = 0.01,
      conf = 0.6,
      minlen = 2,
      maxlen = 4
    )
  )

quality(rules)$improvement <- interestMeasure(rules, measure = "improvement")
rules <- sort(rules, by = "improvement")
inspect(rules[1:10])
```

```{r}
# Finding redundancy
redundant <- is.redundant(rules)
rules_pruned <- rules[!redundant]
rules_pruned <- sort(rules_pruned, by = "confidence")
inspect(rules_pruned[1:10])
```

Having a more relaxed set of rules to get products that go with the most frequently redeemed coupons
```{r}
coupon_rules <-
  apriori(transaction,
    parameter = list(
      supp = 0.003,
      conf = 0.1,
      minlen = 2,
      maxlen = 3
    )
  )

quality(coupon_rules)$improvement <-
  interestMeasure(coupon_rules, measure = "improvement")
coupon_rules <- sort(coupon_rules, by = "improvement")
```

```{r}
# inspectDT(coupon_rules)
```

This is how you read the results,
IF the items in LHS are purchased, THEN the item in RHS is purchased
